Eddy covariance towers as sentinels of abnormal radioactive material releases

### 2024-08-29 manuscript/manuscript.tex
text edits

A potential alternative to active monitoring is passive monitoring \citep[e.g., the German Integrated Measuring and Information system and the International Monitoring System,][]{bieringerMonitoringGroundlevelAir2004}. Although existing wide area monitoring systems are already critical components of abnormal release detection and tracking programs \citep{mediciIMSRadionuclideNetwork2001}, diversification-of-approach may increase the programs' sensitivity and comprehensiveness. With a greater diversity of methods, monitoring programs may be able to detect a wider array of constituents with differing characteristics or expand their geographic extent without the need to deploy new instrumentation. In the present study, we developed a novel passive monitoring approach to supplement existing programs that involves the use of eddy covariance data to capture a signal from known abnormal releases. This signal is then used to investigate whether additional potential abnormal releases exist at other times in the data record. We focused on eddy flux towers, as they have to potential to provide abnormal release signatures, are globally distributed, have been in near continuous operation for several decades, and can be relatively cheaply and easily deployed in locations of interest. Therefore, they constitute a rich target for data mining approaches aimed at identifying signatures resulting from direct interference of radioactive material with target measurements and/or the indirect signal of vegetation responses to the deposition of radioactive material.

@@

We specifically focused on short-term responses of vegetation (on the order of days to weeks) to potential deposition of radioactive material rather than chronic long-term signals (on the order of years to decades). In the first case study, we tested the ability of various data mining techniques to recover the signal from the closest tower (approximately 18 km away) to a known release coming from the Institut des Radioelements (IRE) in Belgium (hereafter called the known release location or the “facility”) nominally occurring on August 23, 2008. The IRE release originated from waste tanks into the atmosphere and lasted for several days totaling 50 GBq 131I \citep{carle2010individual}. For example, grass samples collected in the vicinity of the facility found radioactivity levels of 5000 Bq/kg. In the second case study, we further explored potential signals influx tower data related to the Fukushima Nuclear Disaster nominally occurring on March 11, 2011 releasing a massive radioactive plume that reached North America in 5 days that eventually dispersed throughout the entire northern hemisphere \citep{meszarosPredictabilityDispersionFukushimaderived2016}. We used the signal identified during the known release to identify additional potential abnormal releases occurring at other times in the data record and at other nearby eddy flux towers. While the body of literature on direct or indirect effects of radioactive material on vegetation is limited, we expect that exposure may cause declines in photosynthetic $CO_2$ uptake. This may be expressed as a change in the slope of the $CO_2$ relationship with other environmental variables (e.g. air temperature, humidity, precipitation, etc.) before, during, and after exposure.

@@

For the IRE case study, we gathered measurements from eddy covariance platforms (i.e., “flux towers”) located in Belgium, which are part of the ICOS (Integrated Carbon Observation System) network available through the European Fluxes Database (http://www.europe-fluxdata.eu/). Three sites in proximity to the known release location had sufficient long-term data for our anticipated data mining efforts (BE-Lon, BE-Bra, and BE-Vie sites, Figure 1). Each of the three flux tower sites are located within a crop ecosystem type and had almost 10 years of data available from 2004 through 2013 recorded at 30 min intervals. While BE-Lon had a near-continuous data record during this period, BE-Bra had a long period of missing data in 2005, and both BE-Bra and BE-Vie had a long period of missing data in 2009. The distance between IRE and the flux tower sites are approximately 19 km, 95 km, and 105 km for the BE-Lon, BE-Bra, and BE-Vie respectively.

@@

\begin{table}[]
	\centering
	\caption{Overview of $CO_2$ vs air temperature interaction models at the IRE site. $R^2$ is the coefficient of determination of a linear model fit to the two covariates for the period of the abnormal release event (corresponding to “during” in Fig. 2). Effect is the event effect size of the interaction term 'during' the known release period matching the quantile of the distribution as specified by event quantile effect hyperparameter. Wind (t\%) is the average percentage of time that wind direction was from the site of known release towards the selected flux tower during the known release event. The remaining columns indicate the base-case hyperparameter settings.}
	\begin{tabular}{@{}lllllll@{}}
	\toprule
	Site & $R^2$ & Effect & Wind(\%) & Days(n) & Wind(tol) & Effect(q) \\ \midrule
	BE-Vie & 0.35 & 36.2  & 34 & 7 & 10 & 0.9  \\
	BE-Bra & 0.4 & 57.59  & 5 & 7 & 10 & 0.9 \\
	BE-Lon & 0.47 & 30.13 & 34 & 7 & 10 & 0.9 \\ \bottomrule
	\end{tabular}	
	\label{table:1}
	\end{table}

@@

\section{Discussion}
Using flux tower records collected closest to the IRE release location, our event detection algorithm was able to identify both a known abnormal release and plausible previously unidentified abnormal events. There was a broad agreement among the closest flux tower and more distant flux towers as to the timing and frequency of these events. This suggests that, at the very least, given a facility known to have a prior abnormal event, flux tower networks are capable of providing a means of passive monitoring to detect subsequent events as well as information on the environmental conditions before, during, and after the event. Furthermore, our approach has promise for situations when there is not a prior known abnormal release. In these cases, it may be possible to continuously mine flux tower records (as they become available) for events that exceed some threshold identified in a more comprehensive cross-site, cross-event benchmarking study.

@@

Another contributor to the success of our approach was the fact that Belgium has a relatively dense network of flux towers with long data records. As a result, we were able to compare event detections at the closest tower against more distant towers (refer to Figure 1), increasing our confidence that we were seeing a true signal and not spurious noise. This level of flux tower density may be found in the United States, Western/Central Europe, Japan, and Australia, which have high densities of flux towers, but not throughout the rest of the world \citep{baldocchi2001fluxnet,pastorello2020fluxnet2015}. As a result, this may impose geographical limits on our approach. Another attribute that likely aided our investigations was the fact that all three of the flux towers we examined were located in similar agricultural ecosystem types. Although investigating whether or not flux tower combinations with differing ecosystems would yield similar results is beyond the scope of the present study, we suspect that the data records among towers in disparate ecosystem types could differ too drastically to be of comparative use. For example, \citet{pastorello2020fluxnet2015} show that towers in the crop ecosystem type, of which all three sites used in this study belong, have a relatively narrow distribution of fluxes (i.e., gross primary production, GPP) compared to towers in the grassland or evergreen broadleaf forest types. One reason for this narrow spread in crop ecosystems may be the relative homogeneity of flux tower footprints in crop ecosystems, which typically extend to within 1000m of the tower depending on atmospheric conditions \citep[e.g., wind speed and direction,][]{chuRepresentativenessEddyCovarianceFlux2021}. Given that towers in other ecosystem types with a more heterogeneous flux tower footprint have a wider distribution of fluxes \citep[e.g.][]{pastorello2020fluxnet2015}, we suspect that having all the towers in our study located in a homogenous ecosystem type was helpful in reducing the spikiness and spread of the data.

@@

The fact that we were able to identify known releases (along with potential unidentified abnormal events) in the data record from both a large event (i.e. Fukushima) as well as a smaller event despite not having a detailed radiological or atmospheric transport model is a strength of our approach. Furthermore, our approach is not limited to a radiological context, any abnormal event that affects the plant community and is reflected in flux tower data is a potential target. Because our approach is light-weight and resistant to systematic errors in the observational record, it has broad applicability beyond specific constituents and ecosystem types to identify a wide variety of limited-duration impacts to the plant community within flux tower footprints to ensure human health and industrial safety.

>>>>

A potential alternative to active monitoring is passive monitoring \citep[e.g., the German Integrated Measuring and Information system and the International Monitoring System,][]{bieringerMonitoringGroundlevelAir2004}. Although existing wide area monitoring systems are already critical components of abnormal release detection and tracking programs \citep{mediciIMSRadionuclideNetwork2001}, diversification-of-approach may increase the programs' sensitivity and comprehensiveness. With a greater diversity of methods, monitoring programs may be able to detect a wider array of constituents with differing characteristics or expand their geographic extent without the need to deploy new instrumentation. In the present study, we developed a novel passive monitoring approach to supplement existing programs that involves the use of eddy covariance data to capture a signal from known abnormal releases. This signal is then used to investigate whether additional potential abnormal releases exist at other times in the data record. We focused on eddy flux towers, as they have the potential to provide abnormal release signatures, are globally distributed, have been in near continuous operation for several decades, and can be relatively cheaply and easily deployed in locations of interest. Therefore, they constitute a rich target for data mining approaches aimed at identifying signatures resulting from direct interference of radioactive material with target measurements and/or the indirect signal of vegetation responses to the deposition of radioactive material.

@@

We specifically focused on short-term responses of vegetation (on the order of days to weeks) to potential deposition of radioactive material rather than chronic long-term signals (on the order of years to decades). In the first case study, we tested the ability of various data mining techniques to recover the signal from the closest tower (approximately 18 km away) to a known release coming from the Institut des Radioelements (IRE) in Belgium (hereafter called the known release location or the “facility”) nominally occurring on August 23, 2008. The IRE release originated from waste tanks into the atmosphere and lasted for several days totaling 50 GBq 131I \citep{carle2010individual}. For example, grass samples collected in the vicinity of the facility found radioactivity levels of 5000 Bq/kg. In the second case study, we further explored potential signals in flux tower data related to the Fukushima Nuclear Disaster nominally occurring on March 11, 2011 releasing a massive radioactive plume that reached North America in 5 days that eventually dispersed throughout the entire northern hemisphere \citep{meszarosPredictabilityDispersionFukushimaderived2016}. We used the signal identified during the known release to identify additional potential abnormal releases occurring at other times in the data record and at other nearby eddy flux towers. While the body of literature on direct or indirect effects of radioactive material on vegetation is limited, we expect that exposure may cause declines in photosynthetic $CO_2$ uptake. This may be expressed as a change in the slope of the $CO_2$ relationship with other environmental variables (e.g. air temperature, humidity, precipitation, etc.) before, during, and after exposure.

@@

For the IRE case study, we gathered measurements from eddy covariance platforms (i.e., “flux towers”) located in Belgium, which are part of the ICOS (Integrated Carbon Observation System) network available through the European Fluxes Database (http://www.europe-fluxdata.eu/). Three sites in proximity to the known release location had sufficient long-term data for our anticipated data mining efforts (BE-Lon, BE-Bra, and BE-Vie sites, Figure 1A). Each of the three flux tower sites are located within a crop ecosystem type and had almost 10 years of data available from 2004 through 2013 recorded at 30 min intervals. While BE-Lon had a near-continuous data record during this period, BE-Bra had a long period of missing data in 2005, and both BE-Bra and BE-Vie had a long period of missing data in 2009. The distance between IRE and the flux tower sites are approximately 19 km, 95 km, and 105 km for the BE-Lon, BE-Bra, and BE-Vie respectively.

@@

\begin{table}[]
	\centering
	\caption{Overview of $CO_2$ vs air temperature interaction models at the IRE site. $R^2$ is the coefficient of determination of a linear model fit to the two covariates for the period of the abnormal release event (corresponding to “during” in Fig. 2). Effect is the event effect size of the interaction term 'during' the known release period matching the quantile of the distribution as specified by event quantile effect hyperparameter. Wind (\%) is the average percentage of time that wind direction was from the site of known release towards the selected flux tower during the known release event. The remaining columns indicate the base-case hyperparameter settings.}
	\begin{tabular}{@{}lllllll@{}}
	\toprule
	Site & $R^2$ & Effect & Wind(\%) & Days(n) & Wind(tol) & Effect(q) \\ \midrule
	BE-Vie & 0.35 & 36.2  & 34 & 7 & 10 & 0.9  \\
	BE-Bra & 0.4 & 57.59  & 5 & 7 & 10 & 0.9 \\
	BE-Lon & 0.47 & 30.13 & 34 & 7 & 10 & 0.9 \\ \bottomrule
	\end{tabular}	
	\label{table:1}
	\end{table}

@@

\section{Discussion}
Using flux tower records collected closest to and in the prevailing wind direction to the IRE and the Fukushima release locations, our event detection algorithm was able to identify both known abnormal releases and plausible previously unidentified abnormal events. There was broad agreement among the closest flux tower and more distant flux towers as to the timing and frequency of these releases, and our approach was supported by the loss of signal at flux towers outside of the prevailing wind direction. This suggests that, at the very least, given a facility known to have a prior abnormal release, flux tower networks are capable of providing a means of passive monitoring to detect subsequent events as well as information on the environmental conditions before, during, and after the event. Furthermore, our approach has promise for situations when there is not a prior known abnormal release. In these cases, it may be possible to continuously mine flux tower records (as they become available) for events that exceed some threshold identified in a more comprehensive cross-site, cross-event benchmarking study.

@@

Another contributor to the success of our approach was our ability to compare event detections at  close towers in prevailing wind directions against more distant towers (refer to Figure 1), increasing our confidence that we were seeing a true signal and not spurious noise. The relatively dense networks of flux towers with long data records leveraged in both locations also aided these case studies. This level of flux tower density may be found in the United States, Western/Central Europe, Japan, and Australia, which have high densities of flux towers, but not throughout the rest of the world \citep{baldocchi2001fluxnet,pastorello2020fluxnet2015}. As a result, this may impose geographical limits on our approach, although installation and maintenance of new flux towers in locations of interest may be cost effective compared to other monitoring approaches. Another attribute that likely aided our investigations was the fact that all three of the flux towers we examined were located in similar agricultural and forested ecosystem types for the IRE and Fukushima case studies respectively. Although investigating whether or not flux tower combinations with differing ecosystems would yield similar results is beyond the scope of the present study, we suspect that the data records among towers in disparate ecosystem types could differ too drastically to be of comparative use. For example, \citet{pastorello2020fluxnet2015} show that towers in the crop ecosystem type, of which all three sites used in this study belong, have a relatively narrow distribution of fluxes (i.e., gross primary production, GPP) compared to towers in the grassland or evergreen broadleaf forest types. One reason for this narrow spread in crop ecosystems may be the relative homogeneity of flux tower footprints in crop ecosystems, which typically extend to within 1000m of the tower depending on atmospheric conditions \citep[e.g., wind speed and direction,][]{chuRepresentativenessEddyCovarianceFlux2021}. Given that towers in other ecosystem types with a more heterogeneous flux tower footprint have a wider distribution of fluxes \citep[e.g.][]{pastorello2020fluxnet2015}, we suspect that having all the towers in the IRE case study located in a homogenous ecosystem type was helpful in reducing the spikiness and spread of the data.

@@

The fact that we were able to identify known releases (along with potential unidentified abnormal events) in the data records from both a large event (i.e. Fukushima) as well as a smaller event despite not having a detailed radiological or atmospheric transport model is a strength of our approach. Furthermore, our approach is not limited to a radiological context, any abnormal event that affects the plant community and is reflected in flux tower data is a potential target. Because our approach is light-weight and resistant to systematic errors in the observational record, it has broad applicability beyond specific constituents and ecosystem types to identify a wide variety of limited-duration impacts to the plant community within flux tower footprints to ensure human health and industrial safety.
